Manuscript

### 2025-03-17 Manuscript_MD_ScoreReLiability.qmd
Update HunterSchmidt Section

In this section, we explore some facets on why we disagree with claims
stated in W&D and H&S. Additionally, we explore some alternative
explanations that might help understand why correcting ES for score
reliability does not lead to an inevitable reduction in heterogeneity.

@@

Hunter & Schmidt claim to have derived an equation which demonstrates
how heterogeneity in score reliability inflates heterogeneity in ESs.
“If the level of reliability is independent of the true effect size
across studies, then, to a close approximation:” (p. 309)

@@

We adjusted @eq-HS using the notation used throughout this text. Here,
$\delta_0$ refers to the ES, undistorted by sampling error but not
corrected for measuring error, $\delta$ refers to the ES, undistorted by
sampling error and measurement error and ρxx‘ refers to score
reliability.

@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
function where heterogeneity and mean value of $\delta$ are weighted by
mean score reliability $\rho_{xx‘}$ and its heterogeneity. While it may
not be self-evident in the equation itself, it is easy to construct
cases where heterogeneity in $\delta_0$ is inflated or deflated by
heterogeneity in $\rho_{xx‘}$, contradicting Hunter & Schmidt’s
interpretation of this equation. In @fig-hs, the results of employing
@eq-HS to compute the heterogeneity (variance) in corrected ES is
highlighted. All else held equal, only the individual level of mean
reliability $\mu_{\rho_{xx‘}}$ is varied from .5 to .8. Mean corrected
ES $\mu_{\delta}$ is held constant at 1, its variance $\tau^2_{\delta}$
held constant at .04 and the variance of score reliability
$\tau^2_{\rho_{xx‘}}$ held constant at .02.

```{r, echo = FALSE}
#| label: fig-HS
#| fig-cap: Implications by Hunter & Schmidt's @eg-HS

mu_delta <- 1
var_delta <- .2^2
mu_alpha <- seq(from = .5, to = .8, length.out = 100)
var_alpha <- .14^2

var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha

ggplot() +
  geom_hline(yintercept = .04, colour = "red", linewidth = 1, linetype = "dashed") +
  geom_line(aes(y = var_delta0, x = mu_alpha), linewidth = 2) +
  theme(panel.background = element_rect(fill = "transparent"), 
        plot.background = element_rect(fill = "transparent", colour = "transparent"), 
        panel.grid.major.y = element_line(colour = "grey"),
        panel.grid.major.x = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "transparent"),
        axis.ticks = element_line(colour = "grey"),
        strip.background = element_rect(fill = "transparent")) +
  labs(y = expression(tau^2~"["~delta[0]~"]"),
       x = expression(mu~"["~rho[xx]~"]"))
```

In @fig-hs, the red line corresponds to the variance of corrected ES.

If everything else is held constant, the variance in uncorrected ES can
be a function only of the mean level of score reliability
$\mu_{\rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
demonstrates that correcting for imperfect score reliability does not
inevitably lead to lower heterogeneity, as the line representing
heterogeneity in uncorrected ES does cross the red line, representing
the heterogeneity in corrected ES. Depending on the mean level of score
reliability, heterogeneity in uncorrected ES can be higher or lower than
heterogeneity in corrected ES.

However, @eq-hs is additionally flawed, as it violates a crucial
definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
level of reliability is independent of the true effect size across
studies \[…\]” (2014, p. 309). However, both parameters, ES $\delta$ and
score reliability ρxx‘, can not be independent parameters by definition.
ES $\delta$ is the “true” standardized mean difference, unaffected by
sampling or measurement error. The ES $\delta_0$ is standardized using
the total standard deviation, which is inflated by measurement error.
Equation 1, demonstrating how Cohen’s d is computed demonstrates how the
mean difference is standardized in this case. As demonstrated further
above in @eg-dct, ES $\delta$ is standardized using the true standard
deviation in the sense of CTT. The fact that the true standard deviation
parameter $\sigma_T$ is found both in @eq-dct and @eq-rel demonstrates
that ES $\delta$ and score reliability $\rho_{xx‘}$ can not be
independent variables. Therefore, the basic premise of @eq-hs is not
fulfilled, as the assumption of variable independence could not be
fulfilled.

>>>>

In the previous section we demonstrated that assumptions formulated in
the literature regarding meta-scientific methodology are false. We found
a large number of phenomena where, contrary to popular claims, score
reliability does not explain ES heterogeneity. Over the following
paragraphs, we discuss how the true relationship between these two
parameters might look like. Therein, we also identify under which
circumstances an attenuation correction can actually reduce ES
heterogeneity - and where it is more likely to inflate it.

@@

Hunter & Schmidt claim to have derived an equation which demonstrates
how heterogeneity in score reliability inflates heterogeneity in ESs.
“If the level of reliability is independent of the true effect size
across studies, then, to a close approximation:” (p. 309). However, as
can be easily verified, the equation presented in Hunter & Schmidt
(ref.), Equation X on p. 309, does not imply that assumption across all
conditions.

@@

We adjusted @eq-HS using the notation used throughout this text. Here,
$\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
undistorted by sampling error but not corrected for measuring error. On
the other hand $\delta$, no index, refers to the corrected ES,
undistorted by sampling error and measurement error. Lastly,
$\rho_{XX'}$ refers to score reliability.

@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
function in which heterogeneity and mean value of $\delta$ are weighted
by mean score reliability $\rho_{xx‘}$ and its heterogeneity. Using
selected values for the parameters in @eq-HS, we demonstrate how the
heterogeneity in standardized ES can be both inflated and deflated by
the presence of differences in score reliability. In @fig-hs, the
results of employing @eq-HS to compute the heterogeneity (variance) in
uncorrected ES $\delta_0$ is highlighted. All else held equal, only the
individual level of mean reliability $\mu_{\rho_{xx‘}}$ is varied from
.5 to .8. Mean corrected ES $\mu_{\delta_0}$ is held constant at 1, its
heterogeneity $\tau_{\delta_0}$ held constant at .2 and the
heterogeneity of score reliability $\tau_{\rho_{xx‘}}$ held constant at
.12.

```{r, echo = FALSE}
#| label: fig-HS
#| fig-cap: Implications by Hunter & Schmidt's @eg-HS \n Note: The red line corresponds to the variance of uncorrected ES $\delta_0$.

library(ggplot2)

mu_delta0 <- 1
var_delta0 <- .2^2
mu_alpha <- seq(from = .5, to = .8, length.out = 100)
var_alpha <- .12^2
mu_delta <- mu_delta0/sqrt(mu_alpha)

# var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha
var_delta <- (var_delta0 - mu_delta^2 * var_alpha) / (mu_alpha^2)

ggplot() +
  geom_hline(yintercept = .04, colour = "red", linewidth = 1, linetype = "dashed") +
  geom_line(aes(y = var_delta, x = mu_alpha), linewidth = 2) +
  theme(panel.background = element_rect(fill = "transparent"), 
        plot.background = element_rect(fill = "transparent", colour = "transparent"), 
        panel.grid.major.y = element_line(colour = "grey"),
        panel.grid.major.x = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "transparent"),
        axis.ticks = element_line(colour = "grey"),
        strip.background = element_rect(fill = "transparent")) +
  labs(y = expression(tau^2~"["~delta~"]"),
       x = expression(mu~"["~rho[xx]~"]"))
```

If everything else is held constant, the variance in corrected ES
$\delta$ can be a function only of the mean level of score reliability
$\mu_{\rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
demonstrates that correcting for imperfect score reliability does not
inevitably lead to lower heterogeneity, as the line representing
heterogeneity in corrected ES does cross the red line, representing the
heterogeneity in uncorrected ES. Depending on the mean level of score
reliability, heterogeneity in uncorrected ES can be higher or lower than
heterogeneity in corrected ES.

However, @eq-hs is additionally flawed, as it violates a crucial
definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
level of reliability is independent of the true effect size across
studies \[…\]” (2014, p. 309). However, if we assume that true score
variance $\sigma^2_T$ varies across administrations, it is unlikely that
this assumption can hold. As demonstrated further above in @eg-dct, ES
$\delta$ is standardized using the true standard deviation $\sigma_T$ in
the sense of CTT. The fact that this parameter is found both in @eq-dct
and @eq-rel demonstrates that ES $\delta$ and score reliability
$\rho_{xx‘}$ are unlikely to constitute truly independent variables, as
long as true score variance is not identical across administrations.
